---
title: "r5r_update"
output: html_document
date: "2023-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## setup library and working directory

```{r}
library(r5r)
options(java.parameters = "-Xmx2G")
library(osmextract)
library(opentripplanner)
library(sf)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tmap)
library(tigris)
library(interp)

setwd("C:/study/R_experiment")
```


# fetching OSM pbf data for two counties 

```{r, eval=T}
SLO_file <- oe_match("Ohio")#,provider = "bbbike")

SLO_streets <- oe_read(SLO_file$url, 
                           provider = "openstreetmap_fr", 
                           download_directory = "state", 
                           layer= "lines", 
                           quiet = TRUE) %>%
  filter(!is.na(highway))

street_clip <- SLO_streets[place,] 

tmap_mode("plot")
tm_shape(SLO_streets) + tm_lines()
```

```{r}
graph_path <- c("C:/study/R_experiment/state")
r5r_core <- setup_r5(graph_path, verbose = FALSE)
```

```{r}
x <- "ttt"
print(paste("I need:",x))

grand_ttm$process_county[grand_ttm$County == "Auglaize County"|grand_ttm$County == "Champaign County"|grand_ttm$County == "Hardin County"|grand_ttm$County == "Logan County"|grand_ttm$County == "Shelby County"| grand_ttm$County == "Union County"] <- "Logan"

```


# a for loop for other counties car ttm

```{r}
# "Williams","Fulton", "Defiance", "Henry", "Logan", "Paulding", "Putnam", "Wood", "Hancock",  "Van Wert", "Allen", "Hardin","Mercer","Auglaize", "Union", "Champaign","Shelby", "Darke", "Miami","Clark","Madison","Preble", "Greene","Fayette","Warren","Clinton","Clermont","Highland","Brown","Adams")

#"Williams","Fulton", "Defiance", "Henry", "Logan", "Paulding", "Putnam", "Wood", "Hancock",  "Van Wert",
                 #"Allen"
county_list <- c( "Hardin","Mercer","Auglaize", "Union", "Champaign","Shelby", "Darke", "Miami","Clark",
                 "Madison","Preble","Greene","Fayette","Warren","Clinton","Clermont","Highland","Brown","Adams",
                 "Ottawa","Sandusky","Seneca","Marion","Delaware","Ross","Pike","Scioto","Huron","Crawford",
                 "Morrow","Richland","Knox","Licking","Fairfield","Hocking","Vinton","Jackson","Lawrence","Lorain",
                 "Lake","Ashtabula","Geauga","Medina","Ashland","Wayne","Columbiana","Carroll","Jefferson","Harrison",
                 "Tuscarawas","Holmes","Coshocton","Muskingum","Guernsey","Belmont","Monroe","Noble","Morgan","Perry",
                 "Washington","Meigs","Gallia")

for(j in 1:length(county_list)){
  county <- county_list[j]
  
  CT_center = read.csv("C:/study/public health/CDC/population_weighted_CT_center.csv", as.is=T) %>%  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326) #%>% st_transform(26917)  #all census tract center
  tracts <- tracts("ohio") %>% st_transform(4326) #all ohio tract polygon
  CT_center <- st_join(CT_center, tracts, join=st_within, left=T) %>% rename(id=GEOID) #centroids joined with polygon for GEOID
  # filtering specific county boundary
  place <- counties("ohio") %>%
    st_transform(4326) %>%
    filter(NAME==county) %>%
    mutate(id=GEOID)
  place_buf <- st_buffer(place, 20000, endCapStyle = "FLAT") #20 km buffer around the county
  pop_center <- CT_center[place,] #selected centers within the county
  tract_sel <- tracts[place,] %>% rename(id=GEOID) #selected tracts within the county
  
  updated_data = read.csv("cdc_vax_providers.csv", as.is=T) %>% rename(id=provider_location_guid) #all vax location
  vax_loc = st_as_sf(updated_data, coords = c("longitude", "latitude"), crs = 4326) #%>% st_transform(26917)
  vac_loc <- vax_loc[place_buf,] #vax loc of the county
  vac_loc <- vac_loc %>% distinct(id, .keep_all = TRUE)
  
  
  for(i in 1:nrow(vac_loc)){
    print(paste0("Records processed:" , i))
    ttm_unit = travel_time_matrix(r5r_core = r5r_core,
                            origins = vac_loc[i,],
                            destinations = pop_center,
                            mode = c("CAR"),#TRANSIT","WALK"),#
                            departure_datetime = as.POSIXct("25-05-2022 11:00:00",
                                                            format = "%d-%m-%Y %H:%M:%S"),
                            max_walk_dist = 500,
                            max_trip_duration = 500,
                            verbose = FALSE)
    if(i==1){ttm = ttm_unit} else{
      ttm=rbind(ttm_unit,ttm)}
  }
  summary(ttm$travel_time)
  # percentage of tracts reachable in 45 mins
  below45 <- ttm %>% filter(travel_time<45)
  length(unique(below45$toId))/length(unique(ttm$toId))*100
  
  
  #add facility information
  colnames(ttm)[1] = "id"
  ttm$id <- as.character(ttm$id)
  ttm_county <- merge(ttm,vac_loc,all.x=T, by="id")
  colnames(ttm)[1] = "fromId"
  colnames(ttm_county)[1] = "fromId"
  #add date, mode, county names
  ttm_county <- ttm_county %>% mutate(processing_date = "5/1/23" , mode="car", process_county=county)
  
  grand_ttm <- rbind(grand_ttm, ttm_county)
  print(paste("Processing done for:",county))
}

```


#set county name
```{r}
county <- "Williams"
```

# getting place (county) as destination

```{r}


CT_center = read.csv("C:/study/public health/CDC/population_weighted_CT_center.csv", as.is=T) %>%  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326) #%>% st_transform(26917)  #all census tract center

tracts <- tracts("ohio") %>% st_transform(4326) #all ohio tract polygon

CT_center <- st_join(CT_center, tracts, join=st_within, left=T) %>% rename(id=GEOID) #centroids joined with polygon for GEOID

# filtering specific county boundary
place <- counties("ohio") %>%
  st_transform(4326) %>%
  filter(NAME==county) %>%
  mutate(id=GEOID)

place_buf <- st_buffer(place, 20000, endCapStyle = "FLAT") #20 km buffer around the county

pop_center <- CT_center[place,] #selected centers within the county

tract_sel <- tracts[place,] %>% rename(id=GEOID) #selected tracts within the county

```



# grid analytics (not needed)

```{r, eval=FALSE}
#making grids on the place
grid <- st_sf(st_make_grid(place, 
                           square = T, 
                           cellsize = 0.005,
                           #n = c(500,500),
                           what = "polygons")) %>%
  st_filter(place) 

colnames(grid) <- "geometry"
st_geometry(grid) <- "geometry"

grid <- grid %>%
  mutate(id = seq(1, length(grid$geometry), by=1))

#tm_shape(grid) + tm_borders()

grid_center <- st_centroid(grid)


coords <- st_coordinates(grid_center)
lat <- coords[, 2]
long <- coords[, 1]

# Combine with original data and create new data frame
grid_center <- cbind(grid_center, lat, long)




#tm_shape(grid_center) + tm_dots()

```



# getting origin locations (pediatric locations)

```{r}

updated_data = read.csv("cdc_vax_providers.csv", as.is=T) %>% rename(id=provider_location_guid) #all vax location

vax_loc = st_as_sf(updated_data, coords = c("longitude", "latitude"), crs = 4326) #%>% st_transform(26917)

vac_loc <- vax_loc[place_buf,] #vax loc of the county

vac_loc <- vac_loc %>% distinct(id, .keep_all = TRUE)

```

```{r, eval=F}
# process latlong columns if not present with any sf object
coords <- st_coordinates(vax_loc)
lat <- coords[, 2]
long <- coords[, 1]
vax_loc <- cbind(vax_loc, lat, long)




#ped_lucas = pediatricLocs  %>% filter(County=="Lucas County")

#ped_hamilton <-st_read("C:/study/public health/CDC/shapes/fac_hamilton.shp") %>% rename(id = provider_l)


#tm_shape(ped_cuyahoga) + tm_dots()




#pediatricLocs = rbind(ped_hamilton,ped_cuyahoga) %>%  rename(id = provider_l) # to cpmply with r5r
  #slice(1:2) # removed the duplicates
```



# initiate r5r
```{r, eval=F}
graph_path <- c("C:/study/R_experiment/graphs/")
r5r_core <- setup_r5(paste0(graph_path,county), verbose = FALSE)
```

```{r, echo=F }
 #2021-11-03 12:10:00" # (cuyahoga)"2022-07-11 11:00:05"#(athens) "2017-11-18 11:00:00" #erie "2022-08-24 11:00:00" #(lucas) "2021-09-13 11:00:22" 
# (trumbull)"2022-06-06 11:00:00" #(portage) "2019-12-16 13:54:00"  #(stark) "2022-05-31 11:04:00" # (summit) "2022-07-26 10:10:00" 
#butler "2022-07-13 12:55:00" # montgomery 2022-05-25 17:07:00" (saturday by mistake) # franklin 2021-11-03 12:10:00" #hamilton "2022-06-13 05:15:00" #allen 

#car: cuyahoga, hamilton, franklin, athens,lucas, trumbull, portage, stark,summit, butler, montgomery, allen

for(i in 1:nrow(vac_loc)){
  print(paste0("Records processed:" , i))
  ttm_unit = travel_time_matrix(r5r_core = r5r_core,
                          origins = vac_loc[i,],
                          destinations = pop_center,
                          mode = c("CAR"),#TRANSIT","WALK"),#
                          departure_datetime = as.POSIXct("25-05-2022 11:00:00",
                                                          format = "%d-%m-%Y %H:%M:%S"),
                          max_walk_dist = 500,
                          max_trip_duration = 500,
                          verbose = FALSE)
  if(i==1){ttm = ttm_unit} else{
    ttm=rbind(ttm_unit,ttm)}
}

summary(ttm$travel_time)

# percentage of tracts reachable in 45 mins
below45 <- ttm %>% filter(travel_time<45)

length(unique(below45$toId))/length(unique(ttm$toId))*100



```


```{r, eval=F}


# creating travel time matrix

ttm1_20 <- travel_time_matrix(r5r_core = r5r_core,
                          origins = vac_loc[1:20,],
                          destinations = pop_center,
                          mode = c("TRANSIT","WALK"),
                          departure_datetime = as.POSIXct("13-06-2022 05:15:00",
                                                          format = "%d-%m-%Y %H:%M:%S"),
                          max_walk_dist = 500,
                          max_trip_duration = 500,
                          verbose = FALSE)
```

#fact check
```{r, eval=F}
(ttm)
```


```{r, eval=F}
#select some vaccine location
vac_sel <- vac_loc %>% filter(id=="e9b46bff-c42c-47b7-ba94-ec00c0345f9d")

#select some population center
pop_sub <- pop_center %>% filter(id=="39113000100"|id=="39113000200"|id=="39113000700")

#select matched entries from ttm
ttm_sel <- grand_ttm %>% 
  filter(mode=="car") %>%
  filter(toId=="39113000100"|toId=="39113000200"|toId=="39113000700")  %>%
  filter(fromId=="e9b46bff-c42c-47b7-ba94-ec00c0345f9d")

# record the travel time
ttm_sel

#find the locations and check from google map
tmap_mode("view")

tm_shape(pop_sub) +tm_dots() + tm_text("id") +
  tm_shape(vac_sel) +tm_dots("red") + tm_text("Address") 
```



# process ttm for storage
```{r}

#vac_uniq <- vac_loc %>% distinct(id, .keep_all = TRUE)
#ttm_uniq <- ttm %>% distinct()

#add facility information
colnames(ttm)[1] = "id"
ttm$id <- as.character(ttm$id)
ttm_county <- merge(ttm,vac_loc,all.x=T, by="id")

colnames(ttm)[1] = "fromId"
colnames(ttm_county)[1] = "fromId"

#add tract information
#colnames(ttm_county)[2] = "id" #set it for ttm_county
#ttm_county$id <- as.character(ttm_county$id)
#ttm_county <- left_join(ttm_county,tract_sel, by="id")
#colnames(ttm_county)[2] = "toId"

#add date, mode, county names
ttm_county <- ttm_county %>% mutate(processing_date = "5/1/23" , mode="car", process_county=county)

#from the ttm if I need all tracts of a county, with the status of travel time cutoff, from certain type of facility, created in some date
```

```{r}
grand_ttm <- ttm_county
grand_ttm <- read.csv("grand_ttm.csv", header = T)
```

```{r}
grand_ttm <- rbind(grand_ttm, ttm_county)
```

```{r}
grand_save <- write.csv(grand_ttm, "grand_ttm62C.csv")
```



#build tract subset as isochrones from vac location (change into grid in place of tract_sel as needed)
```{r}
ttm$id <- as.character(ttm$id)
colnames(ttm)[2] = "id"  #change fromID to id to match with grid id

test <- ttm %>%
  merge(tract_sel, by="id", all.x=T) %>% #join grids
  filter(travel_time<10) %>% #filter out other times
  group_by(fromId) %>% #from every vac location
  summarize(geometry = st_union(geometry)) %>% #dissolve all the grids within the time cutoff
  st_as_sf() 
  #summarize(iso = aggregate(id))
 
nrow(test)
tm_shape(test) + tm_polygons("fromId")
```



```{r, eval=F}
# extract OSM network
street_net <- street_network_to_sf(r5r_core)

```

# grid's travel time to the nearest facility.
```{r}

#grid_center_df <- grid_center %>% st_drop_geometry() %>% as.data.frame.table()
colnames(grid_center)[1] <- "toId"

grid_center$toId <- as.character(grid_center$toId)

# add coordinates of destinations to travel time matrix

ttm_j <- left_join(ttm,grid_center, by="toId", copy = F)

#ttm[grid_center_df, on=c('toId' ='id'), `:=`(lon = i.lon, lat = i.lat)]

# interpolate estimates to get spatially smooth result
travel_times.interp <- with(na.omit(ttm_j), interp(long, lat, travel_time, duplicate = 'mean')) %>%
                        with(cbind(travel_time=as.vector(z),  # Column-major order
                                   x=rep(x, times=length(y)),
                                   y=rep(y, each=length(x)))) %>%
                            as.data.frame() %>% na.omit()

# find isochrone's bounding box to crop the map below
bb_x <- c(min(travel_times.interp$x), max(travel_times.interp$x))
bb_y <- c(min(travel_times.interp$y), max(travel_times.interp$y))

#x <- st_as_sf(travel_times.interp, coords = c("x", "y"), crs = 4326)

```

#visualize the minimum travel time
```{r}
# plot
p <- ggplot(travel_times.interp) +
  #geom_sf(data = street_net$edges, color = "gray55", size=0.01, alpha = 0.7) +
  geom_contour_filled(aes(x=x, y=y, z=travel_time), alpha=.7) +
  geom_point(aes(x=long, y=lat, color='Pediatrics location'),
             data=ped_hamilton[c(250,300, 350),]) +
  scale_fill_viridis_d(direction = -1, option = 'B') +
  scale_color_manual(values=c('Pediatrics location'='black')) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  coord_sf(xlim = bb_x, ylim = bb_y) +
  labs(fill = "travel time\n(in minutes)", color='') +
  theme_minimal() +
  theme(axis.title = element_blank())
```
```{r}
tm_shape(x) + tm_dots("travel_time")
```



```{r}
# filtering nearest distance only
near_dist <- ttm  %>%  
  group_by(toId) %>% 
  summarize(near_dist = min(travel_time)) %>%
  replace(is.na(.), 999) %>%
  rename(id = toId) %>%
  left_join(tract_sel, by="id")
st_geometry(near_dist) <- "geometry"  


print(paste0("Number of grid where transit trips to school is not possible: ", nrow(grid) - nrow(near_dist)))


tm_shape(near_dist) + tm_polygons("near_dist") + tm_borders(lwd = 0) +
  tm_shape(ped_hamilton[c(100,200),]) + tm_dots("red", size = 0.05) +
  tm_scale_bar(width = 0.32, position = c(.5,.85)) + 
  tm_compass(size=3,position = c(.8,.5)) + 
  tm_layout(frame = T, title = "Isochrone", 
            title.size = 1.5, title.position = c(.05,.95), legend.outside = T, legend.position = c(.1,.5))


# creating isochrones by dissolving into time cut offs

cut <- list(10,15,20,30,45)
for(i in 1:length(cut)){
  iso <- near_dist %>%
    filter(near_dist<cut[i]) %>%
    st_union() %>%
    st_as_sf() %>%
    mutate(if(i==1){cutoff=paste0("0-",cut[i], " minutes")} else{cutoff=paste0(cut[i-1], "-",cut[i], " minutes")})
  
  if(i==1){all_iso = iso } else{
    all_iso=rbind(iso,all_iso)}
} 

colnames(all_iso)[1]<- "cutoff"


head(all_iso)


tmap_mode("view")
tm_shape(all_iso) + tm_polygons("cutoff", alpha=0.7, style = "jenks",palette = "Spectral", 
                                  title = "Travel time \nIsochrones") + tm_borders("gray", alpha=1, lwd = .2) + tm_shape(place) + tm_borders("black", lwd = .5) +
  tm_shape(place) + tm_borders("black") +
  tm_shape(ped_hamilton[c(100,200),]) + tm_dots("red", size = 0.2) +
  tm_scale_bar(width = 0.32, position = c(.5,.85)) + 
  tm_compass(size=3,position = c(.8,.5)) + 
  tm_layout(frame = T, title = "Isochrone", 
            title.size = 1.5, title.position = c(.05,.95), legend.outside = T, legend.position = c(.1,.5))

hamilton_T <-st_read("C:/study/R_experiment/COVIDPediatricVacAccess/isochoneTransit_hamilton.gpkg") 
tm_shape(hamilton_T) + tm_fill("time", style = "cat")

```

```{r}
tm_shape(hamilton_T) + tm_fill("time", style = "cat")
```

